Distributed Topology Discovery in Self-Assembled Nano Network-On-Chip

### 2014-09-10 trunk/PAPER/disr.tex
cancelled to canceled

A detailed description of the \disr{} execution model as depicted in
Figure~\ref{fig:disr_events} will be discussed in
Subsection~\ref{sub:phases}. Roughly speaking, the execution phases consist of:
\begin{enumerate}
\item Injection of the \disr{} process from an upper layer to set a \emph{bootstrap
node}.
\item Broadcasting and confirmation to create the first segment of the subnet.
\item Parallel segment requests starting from assigned nodes to discover other
segments. 
\item Resulting segment confirmation/cancel packets are processed.
\item Turn prohibition in each confirmed segment.
\end{enumerate}

This first work is explicitly focused on demostrating
the feasibility and scalability of \disr{} in the particular scenario
assumed; for this reason we will not investigate the further degree of
freedom that the last of the above steps introduces, i.e. choosing
\emph{where} actually placing the restriction in the context of the
segment. This choice would striclty depend on the particular routing
strategy that will be used and could result in some interesting
improvements. However, the scope of this work is understading how \disr{} can
implement a distributed topology discovery, mapping defects and
creating a segment partitioning that will be exploited in a second
design phase which includes choosing and implementing a routing strategy. For this
reason, the last two phases will be addressed as one, since we can
safely assume that a random turn (e.g. the first) will be phohibited
once each segment has been confirmed.

@@

\begin{itemize}
\item{\emph{Free}}: a node that has  not been yet considered  by the \disr{} algorithm.
\item{\emph{Bootstrap}}: a node which has been explicitly set as bootstrap node from
an upper layer via. 
\item{\emph{ActiveSearching}}: a node from which a new segments searching process has
been started and not yet cancelled or confirmed. 
\item{\emph{Candidate}}: a node currently candidate to be part of a segment, but not being itself the node from which the searching process was started. 
\item{\emph{CandidateStarting}}: same as above, but the node is currently being considered as candidate
for starting segment. 
\item{\emph{Assigned}}: a node for which the segment has been determined.  
\end{itemize}

The \emph{LED} are a kind of instant snapshot of the execution of \disr{} algorithm, taken in each node,
consisting in the following variables:

\begin{itemize}
\item{\emph{segID}}: a value used to specify the segment to which the
node has been assigned or is currently candidate to be part of.
\item{\emph{visited}}: a boolean value. When \emph{true}
then a \emph{segID} different from \texttt{NULL} specifies the segment 
to which the node has been assigned. 
\item{\emph{tvisited}}: a boolean value. If \emph{true}, the node is
being considered as candidate for a segment, and the \emph{segID} value
specifies the segment id for which the node is candidate. 
\item{\emph{link\_visited[]}}: an array
of values representing information about attached links, that is, the
segID of the segment owning each link. When \texttt{NULL}, the corresponding link has not yet been
assigned.
\item{\emph{link\_tvisited[]}}: an array of
values representing information about attached links, that is, the \emph{segID} of
the segment for which the link is candidate. When \texttt{NULL}, the link is
not candidate for any segment, i.e. it has already been assigned or it is free.
\end{itemize}

>>>>

A detailed description of the \disr{} execution model as depicted in
Figure~\ref{fig:disr_events} will be discussed in
Subsection~\ref{sub:phases}. Roughly speaking, the execution phases consist of:
\begin{enumerate}
\item Injection of the \disr{} process from an upper layer to set a \emph{bootstrap
node}.
\item Broadcasting and confirmation to create the first segment of the subnet.
\item Parallel segment requests starting from assigned nodes to discover other
segments. 
\item Processing of resulting segment confirmation/cancel packets.
\item Turn prohibition in each confirmed segment.
\end{enumerate}

This first work is explicitly focused on demonstrating
the feasibility and scalability of \disr{} in the particular scenario
assumed; for this reason we will not investigate the degree of
freedom that the last of the above steps introduces, i.e. choosing
\emph{where} to place the restriction in the context of a
segment. This choice would strictly depend on the particular routing
strategy that will be used and could result in some interesting
improvements. The scope of this work is understanding how \disr{} can
implement a distributed topology discovery, mapping defects and
creating a segment partitioning that will be exploited in a second
design phase which includes choosing and implementing a routing strategy. 
For this reason, since we can safely assume that a random turn (e.g. 
the first) can be prohibited once each segment has been confirmed, the 
last two execution phases will be addressed as one.

@@

\begin{itemize}
\item{\emph{Free}}: a node that has not been yet considered by the \disr{} algorithm.
\item{\emph{Bootstrap}}: a node which has been explicitly set as bootstrap node from
an upper layer via. 
\item{\emph{ActiveSearching}}: a node from which a new segments searching process has
been started and not yet canceled or confirmed. 
\item{\emph{Candidate}}: a node currently candidate to be part of a segment, but not 
being itself the node from which the searching process was started. 
\item{\emph{CandidateStarting}}: same as above, but with the node considered as candidate
for starting a segment. 
\item{\emph{Assigned}}: a node for which the segment has been determined.
\end{itemize}

\emph{LED} variables are a kind of instant snapshot of the execution of \disr{} algorithm
consisting in the following variables:

\begin{itemize}
\item{\emph{segID}}: a value used to specify the segment to which the
node has been assigned or is currently candidate to be part of.
\item{\emph{visited}}: a boolean value. When \emph{true}
then a \emph{segID} different from \texttt{NULL} specifies the segment 
to which the node has been assigned. 
\item{\emph{tvisited}}: a boolean value. If \emph{true}, the node is
being considered as candidate for a segment, and the \emph{segID} value
specifies the segment ID for which the node is candidate. 
\item{\emph{link\_visited[]}}: an array
of values containing information about attached links (\emph{segID} of the segment 
owning each link). When \texttt{NULL}, the corresponding link has not yet been
assigned.
\item{\emph{link\_tvisited[]}}: an array of
values representing information about attached links (\emph{segID} of
the segment for which the link is candidate). When \texttt{NULL}, the link is
not candidate for any segment, i.e. it has already been assigned or it is free.
\end{itemize}

### 2014-09-10 trunk/PAPER/simulation.tex
Fixed bio for each author and some other typos
cancelled to canceled

In this section we analyze the results in terms of node coverage and
latency with different network sizes, defect rates and bootstrap
injection points. In particular Figure~\ref{fig:results_coverage}
shows node coverage for \disr{} and Reverse Path Forwarding (RPF) tree
based approach (as adapted in \cite{patwardhan_nanoarch06}) respectively. 
While the first aim of \disr{} is not to
reach the optimal coverage, we still can observe quite good
performances as compared to the tree based approach. For low defect
rates, i.e. below 10\%, both approaches reach near ideal coverage for
each of the tested network sizes. Higher defect rates seems to show
more variance for \disr{}. For examplei, at the high 20\% rate \disr{}
ranges from 60\% to 75\% while RPF remains stable at 75\%. However, we
can consider this coverage sensibility at higher defect rates as still acceptable
for this first, not yet optimized version of the proposed approach.
Note that defect rates beyond 25\% lead to many disconnected
regions of nodes that \disr{} currently cannot handle. This threshold can be
intuitively related to the fact that, in the topologies simulated, a
node has (on average) a cardinality of 4 connections to other nodes. This means
that with a defect rate bigger than 25\% a node has at least one of its four
paths as not-working, and this applies (on average) for each node,
leading to higher probability of consecutive disconnected
nodes, which eventually create disconnected regions. For example Figure~\ref{fig:net} shows a 30x30 network with a 25\%
defect rate in which the bottom-left part cannot be reached due to
disconnected regions. Note that the remaining defective nodes belonging to
connected regions are successfully surrounded by \disr{} coverage; in any
case, these defect levels should be considered as worst case scenarios, so the
achieved coverage of $0.5$ is a satisfying result for this first
version of \disr{}. On the other hand, the network size seems to have a
limited impact when defect rate does not introduce too many disconnected
regions. 
\begin{figure*}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.48\textwidth]{pictures/set1.eps} & 
\includegraphics[width=0.48\textwidth]{pictures/coverage.eps} \\
(a) & (b) 
\end{tabular}
\caption{\disr{} (a) and RPF (b) node coverage}
\label{fig:results_coverage}
\end{figure*}

@@

% TODO: long paper
\subsection{Other Optimizations}
Some optimization parameters, which demonstrated to improve the \disr{}
results, have been fixed to some reasonable default values (see below) and are not subject to
further investigations in this paper; once again the focus here is not the
optimal setup of segments, but just demonstrating a working approach. 
These parameters are:
\begin{itemize}
\item \emph{cycle\_links}: max number of retries across the set of links of
each node. While searching for a free link due to an incoming \texttt{SEGMENT\_REQUEST},
the request itself is cancelled after a given number of tries. This
gives to the preceding node on the path the
chance to test a different route instead of waiting indefinitely.
Default value is set to 1, i.e. each link is tried one time.
\item \emph{time to live (TTL)}: when a segment request is cancelled
along a certain path (using a \texttt{SEGMENT\_CANCEL}), a TTL field is
decreased in the request packet, so that requests that have been
denied too much times are canceled, even if free links are
available. This is not only an optimization to shorten
requests' processes, but also solves some critical situations in which
request packets are blocked in routing loops in a particular path.
\item \emph{bootstrap\_timeout}: number of time units that a bootstrap node
should wait before assuming that a livelock in the starting segment
process has occurred. In the worst case, we can imagine that the longest
path required is the one returning to bootstrap node after having
traveled across all the links. So, although this is just an extreme
situation, a good upper limit can be safely be set to $N \times N$.
\item \emph{bootstrap\_immunity}: in order to avoid the failure of the whole \disr{}
setup process, a bootstrap node should not have defective links.
Enabling this optimization, a bootstrap node is immune to defects.
We may think about a pre-bootstrap phase that properly selects (from upper
layer via) a bootstrap node which is tested as properly connected. We
enabled this optimization, however empirical tests have shown us that only
simulations using bootstrap nodes placed on edges would be heavily
affected by similar issues since these nodes start with a lower number of links,
e.g. corner nodes could only have two connected directions, so even a single
defective link could prevent a \texttt{STARTING\_SEGMENT} packet from coming back to the
bootstrap node to close the loop and create the first segment.
\end{itemize}

>>>>

In this section we analyze the results in terms of node coverage and
latency with different network sizes, defect rates and bootstrap
injection points. In particular Figure~\ref{fig:results_coverage}
shows node coverage for \disr{} and Reverse Path Forwarding (RPF) tree
based approach (as adapted in \cite{patwardhan_nanoarch06}) respectively. 
While the first aim of \disr{} is not to
reach the optimal coverage, we still can observe quite good
performances as compared to the tree based approach. For low defect
rates, i.e. below 10\%, both approaches reach near ideal coverage for
each of the tested network sizes. Higher defect rates seems to show
more variance for \disr{}. For example, at the high 20\% rate \disr{}
ranges from 60\% to 75\% while RPF remains stable at 75\%. However, we
can consider this coverage sensibility at higher defect rates as still acceptable
for this first, not yet optimized version of the proposed approach.
Note that defect rates beyond 25\% lead to many disconnected
regions of nodes that \disr{} currently cannot handle. This threshold can be
intuitively related to the fact that, in the topologies simulated, a
node has (on average) a cardinality of 4 connections to other nodes. This means
that with a defect rate bigger than 25\% a node has at least one of its four
paths as not-working, and this applies (on average) for each node,
leading to higher probability of consecutive disconnected
nodes, which eventually create disconnected regions. For example Figure~\ref{fig:net} shows a 30x30 network with a 25\%
defect rate in which the bottom-left part cannot be reached due to
disconnected regions. Note that the remaining defective nodes belonging to
connected regions are successfully surrounded by \disr{} coverage; in any
case, these defect levels should be considered as worst case scenarios, so the
achieved coverage of $0.5$ is a satisfying result for this first
version of \disr{}. On the other hand, the network size seems to have a
limited impact when defect rate does not introduce too many disconnected
regions. 
\begin{figure*}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.48\textwidth]{pictures/set1.eps} & 
\includegraphics[width=0.48\textwidth]{pictures/coverage.eps} \\
(a) & (b) 
\end{tabular}
\caption{\disr{} (a) and RPF (b) node coverage}
\label{fig:results_coverage}
\end{figure*}

@@

% TODO: long paper
\subsection{Other Optimizations}
Some optimization parameters, which demonstrated to improve the \disr{}
results, have been fixed to some reasonable default values (see below) and are not subject to
further investigations in this paper; once again the focus here is not the
optimal setup of segments, but just demonstrating a working approach. 
These parameters are:
\begin{itemize}
\item \emph{cycle\_links}: max number of retries across the set of links of
each node. While searching for a free link due to an incoming \texttt{SEGMENT\_REQUEST},
the request itself is canceled after a given number of tries. This
gives to the preceding node on the path the
chance to test a different route instead of waiting indefinitely.
Default value is set to 1, i.e. each link is tried one time.
\item \emph{time to live (TTL)}: when a segment request is canceled
along a certain path (using a \texttt{SEGMENT\_CANCEL}), a TTL field is
decreased in the request packet, so that requests that have been
denied too much times are canceled, even if free links are
available. This is not only an optimization to shorten
requests' processes, but also solves some critical situations in which
request packets are blocked in routing loops in a particular path.
\item \emph{bootstrap\_timeout}: number of time units that a bootstrap node
should wait before assuming that a livelock in the starting segment
process has occurred. In the worst case, we can imagine that the longest
path required is the one returning to bootstrap node after having
traveled across all the links. So, although this is just an extreme
situation, a good upper limit can be safely be set to $N \times N$.
\item \emph{bootstrap\_immunity}: in order to avoid the failure of the whole \disr{}
setup process, a bootstrap node should not have defective links.
Enabling this optimization, a bootstrap node is immune to defects.
We may think about a pre-bootstrap phase that properly selects (from upper
layer via) a bootstrap node which is tested as properly connected. We
enabled this optimization, however empirical tests have shown us that only
simulations using bootstrap nodes placed on edges would be heavily
affected by similar issues since these nodes start with a lower number of links,
e.g. corner nodes could only have two connected directions, so even a single
defective link could prevent a \texttt{STARTING\_SEGMENT} packet from coming back to the
bootstrap node to close the loop and create the first segment.
\end{itemize}
